---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hgnc <img src='man/figures/logo.svg' align="right" height="139" />

<!-- badges: start -->
<!-- badges: end -->

The [HUGO Gene Nomenclature Committee (HGNC)](https://www.genenames.org/) is a committee of the Human Genome
Organisation (HUGO) that sets the standards for human gene nomenclature.

The HGNC approves a unique and meaningful name for every known human gene, based
on a group of experts.

The goal of `{hgnc}` is to easily download and import HGNC gene data into R.

## Installation

You can install the development version of `{hgnc}` like so:

``` r
# install.packages("remotes")
remotes::install_github("maialab/hgnc")
```

## Usage

To import the latest HGNC data set in tabular format directly into memory as
a tibble do as follows:

```{r}
library(hgnc)

# Date of HGNC last update
last_update()

# Direct URL to the latest archive
(url <- latest_archive_url())

# Import the data set in tidy tabular format
# NB: Multiple-value columns are kept as list-columns
hgnc_dataset <- import_hgnc_dataset_tsv(url)

dplyr::glimpse(hgnc_dataset)
```

If you prefer to download the data set as a file to disk first, you can use
`download_archive()`.

The original data set does not contain the column `hgnc_id2`, which is added as
a convenience by `{hgnc}`; this is because although the HGNC identifiers should
formally contain the prefix `"HGNC:"`, it is often found elsewhere that they are
stripped of this prefix, so the column `hgnc_id2` is also provided whose values
only contain the integer part.
